---
title: " IN PROGRESS --- Cell Membrane Friction - Stick Slip"
author: "Joerg Heintz, April 2017"
output:
  html_document: default
  pdf_document: default
#runtime: shiny
---

```{r setup, include=FALSE}
# !diagnostics off
#library(xlsx)
library(ggplot2)
library(plotly)
library(knitr)
library(dplyr)
library(grid)
library(gridExtra)
library(broom)
library(shiny)
library(manipulate)
library(RColorBrewer)
knitr::opts_chunk$set(echo = FALSE)
opts_knit$set(root.dir = "/Users/joergheintz/Documents/06_Projects/FrictionCellMembranes")
source("CellMem_Functions.R")
```

#Data Fetching 
## Determine Download Parameter AFM txt-files
```{r, echo = TRUE,  eval=TRUE, results = "hide"}
# path to AFM files on the server or a local machine, all files in that folder will be read
        setwd("/Users/joergheintz/Documents/06_Projects/FrictionCellMembranes")
        mypath <- "/Users/joergheintz/Documents/06_Projects/FrictionCellMembranes/02_Data"  
# provide number of lines that need to be skipped, header of AFM file
        fileLineOffset <- 12 

# spring constant alpha
        alpha = 2.93

# provide column numbers
        ColumnsOfInterest <- c(1,7,8) # provide columnames tha need to be read in. predefined is x, y-trace, y-retrace
        lowlimit = 513 #read lines from "lowlimit" to "uplimit" from each AFT txt file
        uplimit = 4096 #read lines from "lowlimit" to "uplimit" from  each AFT txt file

# collect and stores data from the raw AFM files 
        myRawData<-Ingestion(alpha = alpha, ColumnsOfInterest = ColumnsOfInterest, fileLineOffset = fileLineOffset,
        mypath = mypath, lowlimit = lowlimit , uplimit = uplimit)
        
        
```


## Ingestion and Processing

## Data Manipulation for Plotting
```{r, echo = TRUE, eval = TRUE, results="hide"}
# Determining trace or retrace, this file reads the csv file that is generated in the "Ingestion" function
        myfilename<-"CellMembrane_Friction_20170324.csv"
        y_trace__y_retrace<-"y_trace"  # options are y_trace / Y_retrace
      
# Reads from the RawData File a data subset
#. e.g. trace or retrace, and categorical variables from the AFM
        mydata<-PreprocessingIngestedRawData(myfilename = myfilename, y_trace__y_retrace = y_trace__y_retrace)
        
        
# create a cellmembrane color scale
myColors <- brewer.pal(5,"Set1")
names(myColors) <- levels(mydata$system)
colSc <- scale_colour_manual(name = "system",values = myColors)

        
# Case Definitions
        GelPEMvsFullsytem_2nN_50nN <- as.logical(mydata$system == "GelPEM" | mydata$system == "FullSystem")
        GelPEMvsFullsytem_2nN <- as.logical((mydata$system == "GelPEM" | mydata$system == "FullSystem") & mydata$f_on_samp == "2nN")
        GelPEMvsGelvsFullsytem_2nN <- as.logical((mydata$system == "GelPEM" | mydata$system == "FullSystem" | mydata$system == "Gel") &  mydata$f_on_samp == "2nN")
        GelPEMvsGelvsFullsytem_2nN_50nN <- as.logical((mydata$system == "GelPEM" | mydata$system == "FullSystem" | mydata$system == "Gel"))
        GelPEMvsFullsytem_50nN <- as.logical((mydata$system == "GelPEM" | mydata$system == "FullSystem") & mydata$f_on_samp == "50nN")
        GelvsFullsytem_50nN <- as.logical((mydata$system == "Gel" | mydata$system == "FullSystem") & mydata$f_on_samp == "50nN")
        GelvsFullsytem_2nN <- as.logical((mydata$system == "Gel" | mydata$system == "FullSystem") & mydata$f_on_samp == "2nN")

        #mydata<-mydata[GelPEMvsFullsytem_2nN_50nN, ]
        #mydata<-mydata[GelPEMvsFullsytem_2nN, ]
        #mydata<-mydata[GelPEMvsGelvsFullsytem_2nN, ]
        mydata<-mydata[GelPEMvsGelvsFullsytem_2nN_50nN, ]
        #mydata<-mydata[GelPEMvsFullsytem_50nN, ]
        #mydata<-mydata[GelvsFullsytem_2nN, ]
        #mydata<-mydata[GelvsFullsytem_50nN, ]
```


#Data Processing Pipeline
```{r, eval = TRUE, echo = TRUE, fig.width=10, fig.height=18, results = "hide"}
# Pipeline
## Data PreProcessing, mark sticks and slips
        mydata<-StickSlip(mydata, "half_loop")
## Cluster the sticks and slips, provide data.frame and name for the cluster variable
        mydata<-BuildCluster(mydata, "st_sl")
## calc residuals / take out linear effect of x
        mydata<-InsertResidual(mydata)
## Determine max min slope by dymax/dxmax for each peak
        mydata<-MinMaxSlope(mydata)
## Calculate the segmebts (with x, xend=x+dx, y, yend=y+dy 
        mydata<-Segments(mydata)
## Calculate the segmet for fitted valutes (with x, xend=x+dx, y, yend=y+dy 
        mydata<-SegmentsFit(mydata)

## write and arrange columns
        mydata <- mydata %>% select( ID,half_loop, cluster, mSL, y, dy, dymax, x, dx, dxmax,  reg.slope, reg.intercept,
x.fitted, dx.fitted, y.resid, dy.resid, x.fitted.min, x.fitted.max, y.resid.min, y.resid.max, dx.fitted.max, dy.resid.max, slope.fit, alpha, slope,  st_sl, xmin:ymax, f_on_samp, system, trace_d,  line_rate, length, tip, date)
        write.csv(mydata, file = paste(Sys.Date(), "_StickSlip_Cluster_Slope_Segments_", myfilename), row.names = FALSE )

```

```{r, eval = TRUE, echo = FALSE, fig.show = "asis", fig.width=10, fig.height=5, result = "hide"}
# Histogram dy in dependcy of the slope
dyHistogram(mydata, low = -300000, up = 20000000, same_scale = FALSE,  bi =1000, set_x_marker = 0)
```


```{r, eval = TRUE, echo = TRUE, fig.width=15, fig.height=5}
## Histogram Slopes

SlopeHistogram(mydata = mydata, logscale = FALSE, set_x_marker = 0, bi = 500, low = -2000000, up = 2000000, pr = 0, w = 3, h = 1)
```


```{r, eval = TRUE, echo = TRUE, fig.width=10, fig.height=5}
## Plot Slopes
PlotSlope_1(mydata, logscale = TRUE, c = 3)
```


```{r, eval = TRUE, echo = TRUE, fig.width=10, fig.height=4}
## Plot Slopes
PlotSlope_2(mydata, logscale = FALSE, c = 3)
```


## Plotting
```{r, eval = TRUE, echo = TRUE, fig.width=10, fig.height=18}
## Plotting
PlotCol_RegLine_Fit(mydata, p =0 )
```


```{r, echo = TRUE, eval = TRUE, fig.width=10, fig.height=18}
PlotColFit(mydata, c = 2, p = 0, xfitted = 0)
```

# Peaks & Segments Plots
```{r, echo = TRUE, eval = TRUE, fig.width=10, fig.height=18}
PlotCol_RegLine(mydata, c=2, p = 0)
```


```{r, eval = TRUE, echo = TRUE, fig.width=6, fig.height=5}
PlotBoxSystemForce_AllinOne(mydata)
```


```{r, eval = TRUE, echo = TRUE, fig.width=13, fig.height=5}
PlotBoxSystemForce(mydata)
```

# Residuals-Plot
```{r, eval = TRUE, echo = TRUE, fig.width=10, fig.height=22}
PlotColLm(mydata, c = 2, p = 0)
```



#Variable Description
1. x in [um], AFM output
2. u.def in [V$ 1\times10^-6 V$ ], AFM output
8. st_sl, [1,-1], 1 = stick = positive slope, slip = -1 negative slope
9. cluster in [integer], one cluster represents one peak, each cluster contains n points that belong to that peak
10. sl_fac [points/peak], measure points pro peak 
3. alpha [nN/V], calibration constant for calculating the friction force f_friction
3. f_friction in [nN]
4. f_on_samp in [nN], normal force applied on sample
    i) 2nN
    ii) 10nN
    iii) 20nN
5. x_speed in [um/s], AFM "slider" speed, the speed is calcuated by: lenght[um] x line rate [Hz=1/s] x 2 (half loops)
    i) 1um/s
    ii) x mm/s
14. line rate [1/sec], frequency of the line scanning, AFM scannes several lines (back and forth) over the same length to collect data, and to determine the trace, retrace values. 
15. length [um], the trace or retrace distance
11. trace_d [1,0], scans in the forward (trace=1, 1-512 Bit) and reverse (retrace=-1, 513-1024 Bit) directions
13. half_loop[#], number half loops per experiement, one half loop is one complete trace or retrace
6. x_scan_dist in [um], max scan length of 10um (512 bits) both directions. 
7. substrates [categorical]
    i) bg = 1st control bare gel
    ii) gel_PEM = 2nd control is gel-PEM with polymers on surface
    iii) LB_on_Si = 3rd control, lipid bilayer on silica
    iv) LB_on_gelPEM = fullsystem cell membrane model, lipid bilayer on gel-PEM. 
12. Tip
    i) blunt
* slope.fit, over all slope of a complete half loop 
* slope, slope of a cluster of points in which each point  pn < pn+1 or pn > pn1. A cluster is line with positive or negative slope. 







